---
title: depigner 0.8.1 
subtitle: ''
summary: ''
author: Corrado
date: '2020-06-26'
slug: depigner-0-8-1
categories:
  - rstats
  - eng
tags:
  - depigner
  - r-package
  - Hmisc
  - rms
  - Telegram
  - progress
  - usethis
authors: [admin]
featured: yes
image:
  caption: '[Logo by Elisa Sovrano](https://www.elisasovrano.it)'
  focal_point: 'center'
bibliography: ["depigner-pkg.bib"]
biblio-style: "apalike"
link-citations: true
output:
  blogdown::html_page:
    df_print: paged
---

I’m very exited to announce the first release of `{depigner}` to
CRAN!

> __Pigna__ [_pìn'n'a_] is the Italian word for pine cone.  In jargon,
it is used to identify a task which is boring, banal, annoying, painful,
frustrating and maybe even with a not so beautiful or rewarding result,
just like the obstinate act of trying to challenge yourself in
extracting pine nuts from a pine cone, provided that, in the end, you
will find at least one inside it.

The `{depigner}` aims to provide some useful functions to be used to
solve small everyday problems of coding or analyzing data with R. The
hope is to provide solutions to that kind of small-little problems which
would be normally solved using quick-and-dirty (ugly and maybe even
wrong) patches.

[![](https://img.shields.io/badge/WEB%20site-click%20me-orange.svg)](https://corradolanera.github.io/depigner/)

# Installation
You can install the released version from
[CRAN](https://cran.r-project.org/) directly calling:

```{r, eval = FALSE}
install.packages("depigner")
```


If you would like to be updated with the last development version 
available, you can install it from it's
[source](https://corradolanera.github.io/depigner/) on
[GitHub](https://github.com/) by calling:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("CorradoLanera/depigner")
```

Next, you can attach it to your session as usual by:

```{r}
library(depigner)
```



# Provided Tools

 Tools Category                         | Function(s)              | Aim                                                                     |
:---------------------------------------|:-------------------------|:------------------------------------------------------------------------|
[Harrell's verse](#harrells-verse-tools)| `tidy_summary()`         | _`pander`-ready_ data frame from `Hmisc::summary()`                     |
&nbsp;                                  | `paired_test_continuous` | Paired test for continuous variable into `Hmisc::summary`               |
&nbsp;                                  | `paired_test_categorical`| Paired test for categorical variable into `Hmisc::summary`              |
&nbsp;                                  | `adjust_p()`             | Adjusts P-values for multiplicity of tests at `tidy_summary()`          |
&nbsp;                                  | `summary_interact()`     | data frame of OR for interaction from `rms::lrm()`                      |
&nbsp;                                  | `htypes()`               | Will be your variables continuous or categorical in `Hmisc::describe()`?|
[Statistical](#statistical-tools)       | `ci2p()`                 | Get P-value form estimation and confidence interval                     |
[Programming](#programming-tools)       | `pb_len()`               | Quick set-up of a `progress::progress_bar()` progress bar               |
&nbsp;                                  | `install_pkg_set()`      | Politely install set of packages (topic-related sets at `?pkg_sets`)    |
[Development](#development-tools)       | `use_ui()`               | Activate `{usethis}` user interface into your own package               |
&nbsp;                                  | `please_install()`       | Politely ask the user to install a package                              |
&nbsp;                                  | `imported_from()`        | List packages imported from a package (which has to be installed)       |
[Telegram](#telegram-tools)             | `start_bot_for_chat()`   | Quick start of a `{telegram.bot}` Telegram's bot                        |
&nbsp;                                  | `send_to_telegram()`     | Unified wrapper to send _someRthing_ to a Telegram chat                 |
&nbsp;                                  | `errors_to_telegram()`   | Divert all your error messages from the console to a Telegram chat      |
[Why not?!](#why-not)                   | `gdp()`                  | Do you have TOO much pignas in your back?! ... try this out ;-)         |



## Harrell's Verse Tools

[Harrell](https://www.fharrell.com/)'s packages [`{Hmisc}`](https://cran.r-project.org/web/packages/Hmisc/index.html) and [`{rms}`](https://cran.r-project.org/web/packages/rms/index.html) are
amazing packages to do statistical analyses, especially for clinical 
data and purposes. If you do not know them, you should.

They are that useful and vast that they gain the label of
["Harrell'verse'"](https://www.nicholas-ollberding.com/post/an-introduction-to-the-harrell-verse-predictive-modeling-using-the-hmisc-and-rms-packages/).

One of the functions I use more is its `summary()` methods, especially 
with the option `method = "reverse"` enabled. It is incredibly useful to
get a "table one" like information.




### __Pigna__
Often, my colleagues and I have faced to the problem of
using the data inside the output of the `summary()` function, and I have
seen very many colorful patches to manage this issue.

`{depigner}`: depigner has a `tidy_summary()` pipable function that 
  produces in output a lovely table already adjusted to be processed by
  `{pander}`!

- **`tidy_summary()`**: produces a data frame from the `summary()`
functions provided by `{Hmisc}` [@R-Hmisc] and `{rms}` [@R-rms]
packages ready to be `pander::pander()`ed [@R-pander].
  
Currently it is tested for method _reverse_ only:

```{r}
library(rms, quietly = TRUE)
  options(datadist = 'dd')
library(survival)
library(pander)

dd <- datadist(iris)
my_summary <- summary(Species ~., data = iris, method = "reverse")
tidy_summary(my_summary) %>% 
  pander()


dd <- datadist(heart)
surv <- Surv(heart$start, heart$stop, heart$event)
f    <- cph(surv ~ age + year + surgery, data = heart)
my_summary <- summary(f)
tidy_summary(my_summary) %>% 
  pander()
```




### __Pigna__
What if you have paired samples into the data you would like to 
  `summary()`zed?

`{depigner}`: provide the required object to pass to the `summary()`
function to perform a paired test, both for continuous and categorical
data with two or more groups.^[Pay particular attention to the
`?paired_test_*` documentation of interest because data should be
provided in some specific way due to the internal management of records
from `{Hmisc}`'s functions.]


- **`paired_test_*()`**: Paired test for categorical/continuous 
  variables to be used in the `summary()` of the `{Hmisc}` [@R-Hmisc]
  package:
```{r}
data(Arthritis)
# categorical -------------------------
## two groups
summary(Treatment ~ Sex,
    data    = Arthritis,
    method  = "reverse",
    test    = TRUE,
    catTest = paired_test_categorical
)
## more than two groups
summary(Improved ~ Sex,
    data    = Arthritis,
    method  = "reverse",
    test    = TRUE,
    catTest = paired_test_categorical
)

# continuous --------------------------
## two groups
summary(Species ~.,
    data    = iris[iris$Species != "setosa",],
    method  = "reverse",
    test    = TRUE,
    conTest = paired_test_continuous
)
## more than two groups
summary(Species ~.,
    data    = iris,
    method  = "reverse",
    test    = TRUE,
    conTest = paired_test_continuous
)
```
  
  


### __Pigna__
How often were you asked to provide "all vs. all" checks and tests to
find that unique randomly sampled P-value, which is less than 0.05 by a
factor of $10^{-15}$ order of magnitude?
`{depigner}`: gives you a function to automatically adjust the resulting
p-values for multiplicity into the `summary()` tables are also provided.

- **`adjust_p()`**: Adjust P-values of a `tidy_summary` objects:
```{r}
my_summary <- summary(Species ~., data = iris, method = "reverse")

  tidy_summary(my_summary) %>%
    adjust_p() %>%
    pander()
```


Producing a lovely, concise, printable table when interaction kicks in
into the model is the aim of the `summary_interact()` function.

- **`summary_interact()`**: Produce a data frame of OR (with the
corresponding CI95%) for the interactions between different combination
of a continuous variable (for which it is possible to define the
reference and the target values) and (every or a selection of levels of)
a categorical one in a logistic model provided by `lrm()` (from the
`{rms}` package [@R-rms]):

```{r, include = FALSE}
data('transplant')
censor_rows <- transplant[['event']] != 'censored' 
transplant <- transplant[censor_rows, , drop = FALSE]

dd <- datadist(transplant)
lrm_mod <- lrm(event ~ rcs(age, 3)*(sex + abo) + rcs(year, 3),
  data = transplant
)
```

```{r}
summary_interact(lrm_mod, age, abo) %>%
  pander()

summary_interact(lrm_mod, age, abo, p = TRUE) %>%
  pander()

```




### __Pigna__
Another super useful function provided by `Hmisc` is `describe()`, which has a lovely `plot()` method for its results. I had faced some issue when I needed to use those plots programmatically because their outputs are not that clear or easy to infer before to see them. I.e., suppose you want to know which variable will be plotted, and in which plot they will be plotted (or even if a plot will be produced). Hence, working with `{Hmisc}`, it is crucial to know which variables Harrell considers continuous or categorical. Among them, it is also mandatory to know which one he considers to have (good) enough information to have been plotted.

`{depigner}` solve those problems with its family of `htypes()` functions, which seems to have "reasonable" results... __officially__:

```{r echo=FALSE}
blogdown::shortcode('tweet', '1268881861746610176')
```

- **`htypes()`** and friends: get/check types of variable with respect 
  to the `{Hmisc}` ecosystem [@R-Hmisc].
  
```{r}
htypes(mtcars)

desc <- Hmisc::describe(mtcars)
htypes(desc)
htype(desc[[1]])
is_hcat(desc[[1]])
is_hcon(desc[[1]])
```




## Statistical Tools

### __Pigna__
Some people can and prefer to read more information into
a (single) value than into two...


`{depigner}`: have a wrapper to provide an answer to them.

- **`ci2p()`**: compute the p-value related with a provided confidence
  interval (assuming every type o necessary assumptions):
```{r}
ci2p(1.125, 0.634,	1.999, log_transform = TRUE)
```




## Programming Tools

### __Pigna__
Progress bars provided by [`{progress}`](https://cran.r-project.org/web/packages/progress/index.html) are great, complete, and powerful. They are also quite easy to set up, but for effortless usage (which comes quite often) an even more easy-to-use wrapper could be useful.^[Note: currently, after [erum2020](https://2020.erum.io/) I am considering to switch to [`{progressr}`] https://github.com/HenrikBengtsson/progressr) package. However, for the moment, I still consider `{progress}` and our `{depigner}` wrapper as simple to use as powerful for most cases.]

`{depigner}`: helps you with this super simple wrapper, you have to know the length (i.e., number of steps) the progress should do overall, and that's it!

- **`pb_len()`**: Progress bar of given length, wrapper from the 
  `{progress}` [@R-progress] package:
```{r, eval=FALSE}
pb <- pb_len(100)

for (i in 1:100) {
    Sys.sleep(0.1)
    tick(pb, paste("i = ", i))
}
```


### __Pigna__
How often have you (re)installed the same set of packages on (your or other) computers? You need to remember which packages you or they need, you can forget one of them, and next... you have to repeat it for the next computer!

`{depigner}`: gives you a function and a set of predefined sets of packages which you can use to install all the packages you need in a  single call. Moreover, it does it politely querying the use if they agree for the installation, and possibly for a general update.

- **`install_pkg_set()`**: Simple and polite wrapper to install sets of
  packages. Moreover, `{depigner}` provides some sets already defined
  for common scenario in R (analyses, production, documenting, ...).
  See them by call `?pgk_sets`.
```{r, eval=FALSE}
install_pkg_set() # this install the whole `?pkg_all`
install_pkg_set(pkg_stan)

?pkg_sets
```




## Development Tools


### __Pigna__
When developing a package, use the [`{usethis}`](https://usethis.r-lib.org/) interface would be fantastic. Anyway, importing `{usethis}` can be not enough: you still need to  write the call for its functions explicitly, e.g., you could be face to  the problem of writing something like

```r
usethis::ui_stop('field {usethis::ui_field("foo")} must has value {usethis::ui_value(value)}, ...')
```

Which is not that nice, doesn't it?

`{depigner}`: help you, in a single call in the `{usethis}` style to  add to your `NAMESPACE` all the `{usethis}`' UI permitting to you to  go straight to something like

```r
ui_stop('field {ui_field("foo")} must has value {ui_value(value)}, ...')
```

- **`use_ui()`**: Use `{usethis}`' user interface [@R-usethis] in your 
  package
```{r, eval=FALSE}
# in the initial setup steps of the development of a package
use_ui()
```


### __Pigna__
When you need to write code that others should run, it is not kind to force the installation or the update of packages in their environments.
  
`{depigner}`: thanks to a prompt given by __He The Hadley__ during a fantastic workshop in NYC, gives you a function to politely ask others to install on __their__ system the packages that __you__ need.

- **`please_install()`**: This is a polite wrapper to
  `install.packages()` inspired (= w/ very minimal modification) by a
  function Hadley showed us during a course.
```{r, eval=FALSE}
a_pkg_i_miss <- setdiff(available.packages(), installed.packages())[[1]]
please_install(a_pkg_i_miss)
```

### __Pigna__
When you install or attach packages, often (quite ever), other packages (i.e., dependencies) were installed/attached too. Whose are they? Do we need to install both those packages explicitly or one of them it is sufficient because the other will be automatically installed as a consequence?

`{depigner}`: has a simple single function to provide you a quick answer to those questions.^[Note: it reports only the packages __imported__ and not the __suggested__ ones.]

- **`imported_from()`**: If you would like to know which packages are
  imported by a package (eg to know which packages are required for its
  installation or either installed during it) you can use this function
```{r}
imported_from("depigner")
```




## Telegram Tools

Telegram... I love it!^[Do not ask me why... It is wonderful! That's it!]

### __Pigna__
Have you ever run some long, very long, super long, extra long, computation... on a server? How many times did you need to come back to your PC, or to your `ssh` + `screen` session to check its state, to check the logs, to see if some errors evilly happened?

`{depigner}`: provide you some super easy to use wrappers on some functionalities of [`{telegram.bot}`](https://cran.r-project.org/web/packages/telegram.bot/index.html) package which will permit you to be notified with custom messages, images, and even errors on your phone directly. You need only to define your bot (instruction in the help page), and you are ready to go!

- **Wrappers to simple use of Telegram's bots**: wrappers from the
  `{telegram.bot}` package [@R-telegram.bot]:
```{r, eval=FALSE}
# Set up a Telegram bot. read `?start_bot_for_chat`
start_bot_for_chat()

# Send something to telegram
send_to_telegram("hello world")

library(ggplot2)
gg <- ggplot(mtcars, aes(x = mpg, y = hp, colour = cyl)) +
    geom_point()
send_to_telegram(
  "following an `mtcars` coloured plot",
  parse_mode = "Markdown"
)
send_to_telegram(gg)

# Divert output errors to the telegram bot
errors_to_telegram()
```




## Why Not?!

### __Pigna__
SSometimes piñatas are too many to be solved, and you just have to give in to them.
  
`{depigner}`: is near you.^[Why "gdp"?... no one knows it...]

- **`gdp()`**: A wrapper to relax.
```{r, eval=FALSE}
gdp(7)
```

# Acknowledgements 
The `{depigner}`'s logo was lovely designed by
[Elisa Sovrano](https://elisasovrano.it).


# References

```{r bib, include=FALSE, eval = TRUE}
# create a bib file for the R packages used in this document
knitr::write_bib(
  x    = c('depigner', 'rms', 'Hmisc', 'telegram.bot', 'progress',
           'usethis', 'pander'),
  file = 'depigner-pkg.bib'
)
```
