---
title: depigner 0.8.3
subtitle: ''
summary: ''
author: [admin]
date: '2020-10-02'
slug: depigner-0-8-1
categories:
  - depigner
tags:
  - r-packages
  - rstats
  - eng
featured: yes
image:
  caption: '[Logo by Elisa Sovrano](https://www.elisasovrano.it)'
  focal_point: 'center'
bibliography: ["depigner-pkg.bib"]
biblio-style: "apalike"
link-citations: true
output:
  blogdown::html_page:
    df_print: paged
---

I’m very exited to announce the second release of `{depigner}` to
CRAN!

> __Pigna__ [_pìn'n'a_] is the Italian word for pine cone.  In jargon,
it is used to identify a task which is boring, banal, annoying, painful,
frustrating and maybe even with a not so beautiful or rewarding result,
just like the obstinate act of trying to challenge yourself in
extracting pine nuts from a pine cone, provided that, in the end, you
will find at least one inside it.

The `{depigner}` aims to provide some useful functions to be used to
solve small everyday problems of coding or analyzing data with R. The
hope is to provide solutions to that kind of small-little problems which
would be normally solved using quick-and-dirty (ugly and maybe even
wrong) patches.

[![](https://img.shields.io/badge/WEB%20site-click%20me-orange.svg)](https://corradolanera.github.io/depigner/)

# Installation
You can install the released version from
[CRAN](https://cran.r-project.org/) directly calling:

```{r, eval = FALSE}
install.packages("depigner")
```


If you would like to be updated with the last development version 
available, you can install it from it's
[source](https://corradolanera.github.io/depigner/) on
[GitHub](https://github.com/) by calling:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("CorradoLanera/depigner")
```

Next, you can attach it to your session as usual by:

```{r}
library(depigner)
```



# Provided Tools

For all the provided tools you can look at the [`{depigner}` website](https://corradolanera.github.io/depigner/), or at the [previous post](https://www.corradolanera.it/2020/06/26/depigner-0-8-1/)


# News from the previous version


# Acknowledgements 
The `{depigner}`'s logo was lovely designed by
[Elisa Sovrano](https://elisasovrano.it).


# References

```{r bib, include=FALSE, eval = TRUE}
# create a bib file for the R packages used in this document
knitr::write_bib(
  x    = c('depigner', 'rms', 'Hmisc', 'telegram.bot', 'progress',
           'usethis', 'pander'),
  file = 'depigner-pkg.bib'
)
```
